---
title: "Sunset-Sunrise Survey"
author: "Jeremy Froidevaux"
date: "26 October 2016"
output: html_document
---
</p>
## List of package required
```{r echo=FALSE}
library(ggplot2)
library(rgdal)
library(scales)
library(ggmap)
library(dplyr)
library(Cairo)
library(maps)
library(mapdata)
library(grid)
library(maptools)
library(plyr)
library(plotly)
library(geosphere)
library(data.table)
library(plotrix)
```
</p>
## Theme used with ggplot2
```{r include=FALSE}
theme_jack <- function (base_size = 12, base_family = "") {
  theme_gray(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(colour = "black"),
      axis.title.x = element_text(colour = "black", size=rel(1.5)),
      axis.title.y = element_text(colour = "black", size=rel(1.5),angle=90,margin=margin(0,10,0,0)),
      panel.background = element_rect(fill="white"),
      panel.grid.minor.y =  element_blank(),
      panel.grid.minor.x =  element_blank(),
      panel.grid.major = element_blank(),
      panel.border = element_blank(), 
      plot.background = element_blank()
    )   
}

theme_set(theme_jack())

```
</p>
## Codes to obtain Figure 1 (maps)
</p>
### Creation of a scale bar & north arrow with R
```{r include=FALSE, fig.width=8,fig.height=10}

create_scale_bar <- function(lon,lat,distance_lon,distance_lat,distance_legend, dist_units = "km"){
    # First rectangle
    bottom_right <- gcDestination(lon = lon, lat = lat, bearing = 90, dist = distance_lon, dist.units = dist_units, model = "WGS84")
    
    topLeft <- gcDestination(lon = lon, lat = lat, bearing = 0, dist = distance_lat, dist.units = dist_units, model = "WGS84")
    rectangle <- cbind(lon=c(lon, lon, bottom_right[1,"long"], bottom_right[1,"long"], lon),
    lat = c(lat, topLeft[1,"lat"], topLeft[1,"lat"],lat, lat))
    rectangle <- data.frame(rectangle, stringsAsFactors = FALSE)
    
    # Second rectangle t right of the first rectangle
    bottom_right2 <- gcDestination(lon = lon, lat = lat, bearing = 90, dist = distance_lon*2, dist.units = dist_units, model = "WGS84")
    rectangle2 <- cbind(lon = c(bottom_right[1,"long"], bottom_right[1,"long"], bottom_right2[1,"long"], bottom_right2[1,"long"], bottom_right[1,"long"]),
    lat=c(lat, topLeft[1,"lat"], topLeft[1,"lat"], lat, lat))
    rectangle2 <- data.frame(rectangle2, stringsAsFactors = FALSE)
    
    # Now let's deal with the text
    on_top <- gcDestination(lon = lon, lat = lat, bearing = 0, dist = distance_legend, dist.units = dist_units, model = "WGS84")
    on_top2 <- on_top3 <- on_top
    on_top2[1,"long"] <- bottom_right[1,"long"]
    on_top3[1,"long"] <- bottom_right2[1,"long"]
    
    legend <- rbind(on_top, on_top2, on_top3)
    legend <- data.frame(cbind(legend, text = c(0, distance_lon, distance_lon*2)), stringsAsFactors = FALSE, row.names = NULL)
    return(list(rectangle = rectangle, rectangle2 = rectangle2, legend = legend))
}

create_orientation_arrow <- function(scale_bar, length, distance = 1, dist_units = "km"){
    lon <- scale_bar$rectangle2[1,1]
    lat <- scale_bar$rectangle2[1,2]
    
    # Bottom point of the arrow
    beg_point <- gcDestination(lon = lon, lat = lat, bearing = 0, dist = distance, dist.units = dist_units, model = "WGS84")
    lon <- beg_point[1,"long"]
    lat <- beg_point[1,"lat"]
    
    # Let us create the endpoint
    on_top <- gcDestination(lon = lon, lat = lat, bearing = 0, dist = length, dist.units = dist_units, model = "WGS84")
    
    left_arrow <- gcDestination(lon = on_top[1,"long"], lat = on_top[1,"lat"], bearing = 225, dist = length/5, dist.units = dist_units, model = "WGS84")
    
    right_arrow <- gcDestination(lon = on_top[1,"long"], lat = on_top[1,"lat"], bearing = 135, dist = length/5, dist.units = dist_units, model = "WGS84")
    
    res <- rbind(
            cbind(x = lon, y = lat, xend = on_top[1,"long"], yend = on_top[1,"lat"]),
            cbind(x = left_arrow[1,"long"], y = left_arrow[1,"lat"], xend = on_top[1,"long"], yend = on_top[1,"lat"]),
            cbind(x = right_arrow[1,"long"], y = right_arrow[1,"lat"], xend = on_top[1,"long"], yend = on_top[1,"lat"]))
    
    res <- as.data.frame(res, stringsAsFactors = FALSE)
    
    # Coordinates from which "N" will be plotted
    coords_n <- cbind(x = lon, y = (lat + on_top[1,"lat"])/2)
    
    return(list(res = res, coords_n = coords_n))
}

scale_bar <- function(lon, lat, distance_lon, distance_lat, distance_legend, dist_unit = "km", rec_fill = "white", rec_colour = "black", rec2_fill = "black", rec2_colour = "black", legend_colour = "black", legend_size = 3, orientation = TRUE, arrow_length = 500, arrow_distance = 300, arrow_north_size = 6){
    the_scale_bar <- create_scale_bar(lon = lon, lat = lat, distance_lon = distance_lon, distance_lat = distance_lat, distance_legend = distance_legend, dist_unit = dist_unit)
    # First rectangle
    rectangle1 <- geom_polygon(data = the_scale_bar$rectangle, aes(x = lon, y = lat), fill = rec_fill, colour = rec_colour)
    
    # Second rectangle
    rectangle2 <- geom_polygon(data = the_scale_bar$rectangle2, aes(x = lon, y = lat), fill = rec2_fill, colour = rec2_colour)
    
    # Legend
    scale_bar_legend <- annotate("text", label = paste(the_scale_bar$legend[,"text"], dist_unit, sep=""), x = the_scale_bar$legend[,"long"], y = the_scale_bar$legend[,"lat"], size = legend_size, colour = legend_colour)
    
    res <- list(rectangle1, rectangle2, scale_bar_legend)
    
    if(orientation){# Add an arrow pointing North
        coords_arrow <- create_orientation_arrow(scale_bar = the_scale_bar, length = arrow_length, distance = arrow_distance, dist_unit = dist_unit)
        arrow <- list(geom_segment(data = coords_arrow$res, aes(x = x, y = y, xend = xend, yend = yend)), annotate("text", label = "N", x = coords_arrow$coords_n[1,"x"], y = coords_arrow$coords_n[1,"y"], size = arrow_north_size, colour = "black"))
        res <- c(res, arrow)
    }
    return(res)
}




```
</p>
### Download the required maps (http://www.gadm.org/)
```{r echo=FALSE}
UKmap<-readRDS("C:/Users/Jeremy/Documents/BCT internship/GBR_adm0.rds")
Imap<-readRDS("C:/Users/Jeremy/Documents/BCT internship/IRL_adm0.rds")
Gmap<-readRDS("C:/Users/Jeremy/Documents/BCT internship/GGY_adm0.rds")
Jmap<-readRDS("C:/Users/Jeremy/Documents/BCT internship/JEY_adm0.rds")
```
</p>
### Plot the maps
```{r echo=FALSE, fig.width=8,fig.height=10,message=FALSE}
#Map a
P0<-ggplot()+geom_polygon(data=UKmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")+coord_map()
P1<-P0+geom_polygon(data=Imap,aes(x=long,y=lat,group=group),fill="white",color="black")
P2<-P1+geom_polygon(data=Gmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")
P3<-P2+geom_polygon(data=Jmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")
P4<-P3+geom_point(data=dB,aes(x=X,y=Y),color="darkorange",fill="red")
P5<- P4 + scale_bar(lon = -10.6, lat = 49.5, 
                         distance_lon = 100, distance_lat = 20, 
                         distance_legend = 40, dist_unit = "km", 
                         arrow_length = 100, arrow_distance = 60, arrow_north_size = 6)
P6<-P5 + theme(panel.grid.minor = element_line(colour = NA), panel.grid.minor = element_line(colour = NA),
             panel.background = element_rect(fill = NA, colour = NA), axis.text.x = element_blank(),
             axis.text.y = element_blank(), axis.ticks.x = element_blank(),
             axis.ticks.y = element_blank(), axis.title = element_blank(),
             rect = element_blank(), 
             plot.margin = unit(0 * c(-1.5, -1.5, -1.5, -1.5), "lines"))
P6

#Map b
P0<-ggplot()+geom_polygon(data=UKmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")+coord_map()
P1<-P0+geom_polygon(data=Imap,aes(x=long,y=lat,group=group),fill="white",color="black")
P2<-P1+geom_polygon(data=Gmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")
P3<-P2+geom_polygon(data=Jmap,aes(x=long,y=lat,group=group),fill="grey10",color="black")
P4<-P3+geom_point(data=obs,aes(x=X,y=Y),color="purple",fill="darkblue")
P5<- P4 + scale_bar(lon = -10.6, lat = 49.5, 
                         distance_lon = 100, distance_lat = 20, 
                         distance_legend = 40, dist_unit = "km", 
                         arrow_length = 100, arrow_distance = 60, arrow_north_size = 6)
P6<-P5 + theme(panel.grid.minor = element_line(colour = NA), panel.grid.minor = element_line(colour = NA),
             panel.background = element_rect(fill = NA, colour = NA), axis.text.x = element_blank(),
             axis.text.y = element_blank(), axis.ticks.x = element_blank(),
             axis.ticks.y = element_blank(), axis.title = element_blank(),
             rect = element_blank(), 
             plot.margin = unit(0 * c(-1.5, -1.5, -1.5, -1.5), "lines"))
P6
```
</p>
## Codes to obtain Figure 2 (histogram)
```{r include=FALSE}
#Histogram a: all data
dt<-read.table("C:/Users/Jeremy/Documents/BCT internship/distance.txt",header=TRUE,sep="\t",dec=".")
setDT(dt)
dt[,distance_hav := distHaversine(matrix(c(X1,Y1),ncol=2),matrix(c(X2,Y2),ncol=2))]
av<-mean(dt$distance_hav)/1000
se<-std.error(dt$distance)
med<-median(dt$distance_hav)/1000
d1<-ggplot(dt,aes(distance_hav/1000))+geom_histogram()+labs(x="Distance (in km)")
d1

#Histogram b: remove dara greater than 85th percentile in the data frame. 
dist85<-dt[dt$distance_hav < quantile(dt$distance_hav, 0.85), ]
d2<-ggplot(dist85,aes(distance_hav/1000))+geom_histogram()+labs(x="Distance (in km)")
d2
length(dt$distance_hav[dt$distance_hav < 251])
```
</p>
## Codes to obtain Figure 3 (interactive pie chart)
```{r include=FALSE}
hab<-read.table("C:/Users/Jeremy/Documents/BCT internship/habitat.txt",header=TRUE,dec=".")
ds <- data.frame(labels = hab$Habitat,
                 values= hab$Number_of_sites)
p9<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
p9<- layout(p9,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"), bgcolor = "transparent"),
            autosize = T, xaxis = list(color = "transparent"), yaxis = list(color = "transparent"))
p9
```
</p>
## Codes to obtain Figures 4 & 5 (bar plots)
```{r echo=FALSE}
#Figure 4
Q<-ggplot(data=obs,aes(x=no_of_surveys,y=obs))+geom_bar(stat="identity",fill="darkblue")
Q1<-Q+labs(x="Number of surveys",y="Number of volunteers")+ylim(c(0,100))
Q2<-Q1+annotate("text",x=1,y=88,label="85")
Q3<-Q2+annotate("text",x=2,y=12,label="9")
Q4<-Q3+annotate("text",x=3,y=8,label="5")
Q5<-Q4+annotate("text",x=4,y=4,label="1")
Q6<-Q5+annotate("text",x=5,y=5,label="2")
Q6
#Figure 5
Q<-ggplot(data=obs,aes(x=no_of_surveys,y=maxno_of_surveyors))+geom_bar(stat="identity",fill="darkblue")
Q1<-Q+labs(x="Number of surveys",y="Number of volunteers")+ylim(c(0,150))
Q2<-Q1+annotate("text",x=1,y=142,label="138")
Q3<-Q2+annotate("text",x=2,y=20,label="16")
Q4<-Q3+annotate("text",x=3,y=20,label="16")
Q5<-Q4+annotate("text",x=4,y=5,label="1")
Q6<-Q5+annotate("text",x=5,y=7,label="3")
Q6
```
</p>
## Codes to obtain Figures 6 & 7 (boxplots)
```{r echo=FALSE}
#Figure 6
date<-read.table("C:/Users/Jeremy/Documents/BCT internship/date.txt",header=TRUE,dec=".")
date$month <- factor(date$month,levels = c("May" ,"June","July","August","September"),ordered = TRUE)
ggplot(date, aes(x=month, y=count)) + geom_boxplot()+labs(x=" ",y="Total number of observations")+expand_limits(x = 0, y = 0)

#Figure 7
date2<-read.table("C:/Users/Jeremy/Documents/BCT internship/date2.txt",header=TRUE,dec=".")
date2$month <- factor(date2$month,levels = c("June","July","August"),ordered = TRUE)
p<-ggplot(date2, aes(x=month, y=count)) + geom_boxplot()+labs(x=" ",y="Number of observations per survey")+expand_limits(y = 0)
p

```
</p>
## Codes to obtain Figure 8 (pie chart)
```{r include=FALSE}
# Pie Chart with Percentages
brec2<-read.table("C:/Users/Jeremy/Documents/BCT internship/batrec2.txt",header=TRUE,dec=".")
library(plotly)
ds <- data.frame(labels = brec2$bat_species,
                 values= brec2$count)
p2<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
p2<- layout(p2,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"),
                          bgcolor = "transparent"),
            autosize = T, xaxis = list(color = "transparent"),axis = list(color = "transparent"))
p2

```
</p> 
## Codes to obtain Tables
```{r echo=FALSE}
 library(knitr)
rel<-read.table("C:/Users/Jeremy/Documents/BCT internship/reliability.txt",header=TRUE,quote="",sep="\t",dec=".")
 kable(rel[,1:3])
```
</p>
## Codes to obtain Figure 9 (pie chart)
```{r include=FALSE}
id<-read.table("C:/Users/Jeremy/Documents/BCT internship/spid.txt",header=TRUE,dec=".")
library(plotly)
ds <- data.frame(labels = id$species_identification,
                 values= id$count)
p4<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
p4<- layout(p4,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"),
                          bgcolor = "transparent"),
            autosize = T, xaxis = list(color = "transparent"),axis = list(color = "transparent"))
p4
```
</p>
## Codes to obtain Figure 10 (horizontal barplot)
```{r echo=FALSE}
bspid<-read.table("C:/Users/Jeremy/Documents/BCT internship/batspecid.txt",header=TRUE,dec=".")
bspid2<- ddply(bspid, "bat_species", mutate, count = count/sum(count))

ggplot(data = bspid2, aes(x = bat_species, y = count, fill = species_identification)) + 
    geom_bar(stat="identity") + coord_flip() + 
    scale_y_continuous(labels = percent_format())+ scale_fill_brewer(palette = 12)+labs(x="",y="Records")
```
</p>
## Codes to otain Figure 11 (pie chart)
```{r include=FALSE}
det<-read.table("C:/Users/Jeremy/Documents/BCT internship/det.txt",header=TRUE,dec=".")
library(plotly)
ds <- data.frame(labels = det$detector_type,
                 values= det$count)
p5<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
p5<- layout(p5,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"),
                          bgcolor = "transparent"),
            autosize = T,xaxis = list(color = "transparent"), yaxis = list(color = "transparent"))
p5
```
</p> 
## Codes to obtain Figure 12 (pie chart)
```{r include=FALSE}
bbh<-read.table("C:/Users/Jeremy/Documents/BCT internship/batbehav.txt",header=TRUE,dec=".")
library(plotly)

ds <- data.frame(labels = bbh$type_of_sighting,
                 values= bbh$count)
p3<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
p3<- layout(p3,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"),
                          bgcolor = "transparent"),
            autosize = T,xaxis = list(color = "transparent"), yaxis = list(color = "transparent"))
p3
```
</p> 
## Codes top obtain Figure 12 (horizontal barplot)
```{r echo=FALSE}
bid<-read.table("C:/Users/Jeremy/Documents/BCT internship/behavid.txt",header=TRUE,dec=".")
bid2<- ddply(bid, "type_of_sighting", mutate, count = count/sum(count))

ggplot(data = bid2, aes(x = type_of_sighting, y = count, fill = species_identification)) + 
    geom_bar(stat="identity") + coord_flip() + 
    scale_y_continuous(labels = percent_format())+ scale_fill_brewer(palette = 12)+labs(x="Type of sighting",y="% of records")

```
</p>
## Codes to obtain Figure 14 (pie chart)
```{r include=FALSE}
tax2<-read.table("C:/Users/Jeremy/Documents/BCT internship/taxa2.txt",header=TRUE,dec=".")
library(plotly)
ds <- data.frame(labels = tax2$taxon_group,
                 values= tax2$count)
t1<-plot_ly(labels = ds$labels, values = ds$values, type = "pie")
t1<- layout(t1,
            paper_bgcolor="rgb(255,255,255)",
            plot_bgcolor="rgb(255,255,255)",
            legend = list(font = list(color = "black"),
                          bgcolor = "transparent"),
            autosize = T,xaxis = list(color = "transparent"), yaxis = list(color = "transparent"))
t1
```
</p>
## Codes to obtain Figure 15 & 16 (barplots)
```{r echo=FALSE}
#Figure 15
prev<-read.table("C:/Users/Jeremy/Documents/BCT internship/previous.txt",header=TRUE,quote="",sep="\t",dec=".")
Q<-ggplot(data=prev,aes(x=year,y=no_of_volunteers))+geom_bar(stat="identity",fill="darkblue")
Q1<-Q+labs(x="",y="Number of volunteers")+ylim(c(0,400))
Q2<-Q1+annotate("text",x=2002,y=86,label="66")
Q3<-Q2+annotate("text",x=2003,y=221,label="201")
Q4<-Q3+annotate("text",x=2004,y=272,label="252")
Q5<-Q4+annotate("text",x=2005,y=346,label="326")
Q6<-Q5+annotate("text",x=2006,y=239,label="219")
Q7<-Q6+annotate("text",x=2007,y=213,label="193")
Q8<-Q7+annotate("text",x=2008,y=207,label="187")
Q9<-Q8+annotate("text",x=2009,y=193,label="173")
Q10<-Q9+annotate("text",x=2010,y=223,label="203")
Q11<-Q10+annotate("text",x=2011,y=306,label="286")
Q12<-Q11+annotate("text",x=2012,y=210,label="190")
Q13<-Q12+annotate("text",x=2013,y=197,label="177")
Q14<-Q13+annotate("text",x=2014,y=206,label="186")
Q15<-Q14+annotate("text",x=2015,y=245,label="225")
Q16<-Q15+annotate("text",x=2016,y=122,label="102")
Q17<-Q16+ geom_hline(yintercept=206,linetype=3,colour="red")
Q17

#Figure 16
prev<-read.table("C:/Users/Jeremy/Documents/BCT internship/previous2.txt",header=TRUE,quote="",sep="\t",dec=".")
Q<-ggplot(data=prev,aes(x=year,y=no_of_volunteers))+geom_bar(stat="identity",fill="darkblue")
Q1<-Q+labs(x="",y="Number of volunteers")+ylim(c(0,400))
Q2<-Q1+annotate("text",x=2002,y=111,label="91")
Q3<-Q2+annotate("text",x=2003,y=294,label="274")
Q4<-Q3+annotate("text",x=2004,y=315,label="295")
Q5<-Q4+annotate("text",x=2005,y=377,label="357")
Q6<-Q5+annotate("text",x=2006,y=256,label="236")
Q7<-Q6+annotate("text",x=2007,y=201,label="181")
Q8<-Q7+annotate("text",x=2008,y=292,label="272")
Q9<-Q8+annotate("text",x=2009,y=157,label="137")
Q10<-Q9+annotate("text",x=2010,y=232,label="212")
Q11<-Q10+annotate("text",x=2011,y=272,label="252")
Q12<-Q11+annotate("text",x=2012,y=295,label="275")
Q13<-Q12+annotate("text",x=2013,y=222,label="202")
Q14<-Q13+annotate("text",x=2014,y=219,label="199")
Q15<-Q14+annotate("text",x=2015,y=292,label="272")
Q16<-Q15+annotate("text",x=2016,y=194,label="174")
Q17<-Q16+ geom_hline(yintercept=232.5,linetype=3,colour="red")
Q17
```
</p>
## Codes to obtain Figure 17 (bar plot)
```{r echo=FALSE}
year = c(2015,2016)
hour = c(45,2.3)
df = data.frame(year, hour)
Q<-ggplot(data=df,aes(x=year,y=hour),width=0.1)+geom_bar(stat="identity",fill="darkblue")
Q1<-Q+labs(x="",y="No. of hours")+ylim(c(0,50))+scale_x_continuous(breaks=pretty_breaks(n=2))
Q2<-Q1+annotate("text",x=2015,y=47,label="45.0")
Q3<-Q2+annotate("text",x=2016,y=4,label="2.3")
Q3
```
</p>



